### 2016-10-17 paper.tex
paper

\documentclass[a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{pdfpages}
\usepackage{float, algorithmic, algorithm2e, program}
\usepackage{tabularx}
\usepackage{longtable}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}[section]

\title{Bathtub Qualitative Reasoning}
\author{Nicola De Cao, Luca Falorsi, Govert Verkes}
\date{\today}

\begin{document}
\maketitle

\section{Problem}

\begin{figure}[H]
\includegraphics[]{problem.png}
\caption{Drawings of the causal model active for the system}
\end{figure}

\begin{algorithm}
\If{condition}{then-block}
\caption{Drawings of the causal model active for the system}
\end{algorithm}

\end{document}

### 2016-10-17 paper.tex
paper

@@
 
+\begin{itemize}
+\item \textbf{Quantities:}
+\begin{itemize}
+\item Inflow (of water into the container) $\in [0, +]$
+\item Outflow (of water out of the container) $\in [0, +, Max]$
+\item Volume (of the water in the container) $\in [0, +, Max]$
+\end{itemize}
+
+\item \textbf{Dependencies:}
+
+\begin{itemize}
+\item I+(Inflow, Volume): the amount of inflow increases the volume
+\item I-(Outflow, Volume): the amount of outflow decreases the volume
+\item P+(Volume, Outflow): outflow changes are proportional to volume changes
+\item V(Volume(Max), Outflow(Max)): the outflow is at its highest value (Max), when the volume is at it highest value
+\item V(Volume(0), Outflow(0)): there is no outflow, when there is no volume
+\end{itemize}
+\end{itemize}
+
 \begin{figure}[H]

### 2016-10-20 paper.tex
plt
update

@@
 \usepackage{longtable}
+\usepackage{epstopdf}
 \newtheorem{theorem}{Theorem}
@@
 
-\section{Problem}
+\section{Base Problem}
 
@@
 \begin{figure}[H]
-\includegraphics[]{problem.png}
-\caption{Drawings of the causal model active for the system}
+\centering
+\includegraphics[scale=0.5]{problem1.eps}
+\caption{Drawings of the causal model active for the problem}
 \end{figure}
+
+\section{Algorithm}
 
@@
 \If{condition}{then-block}
-\caption{Drawings of the causal model active for the system}
+\caption{}
 \end{algorithm}
 
+\section{State-graph}
+
+\section{Extra Problem}
+
+\begin{itemize}
+\item \textbf{Quantities:}
+\begin{itemize}
+\item Inflow (of water into the container) $\in [0, +]$
+\item Outflow (of water out of the container) $\in [0, +, Max]$
+\item Volume (of the water in the container) $\in [0, +, Max]$
+\item Height (of the water column in the container) $\in [0, +, Max]$
+\item Pressure (of the water column at the bottom of the container) $\in [0, +, Max]$
+\end{itemize}
+
+\item \textbf{Dependencies:}
+
+\begin{itemize}
+\item I+(Inflow, Volume): the amount of inflow increases the volume
+\item I-(Outflow, Volume): the amount of outflow decreases the volume
+\item P+(Volume, Height): height changes are proportional to volume changes
+\item P+(Height, Pressure): pressure changes are proportional to height changes
+\item P+(Pressure, Outflow): outflow changes are proportional to pressure changes
+\item V(Volume(Max), Height(Max)): the height is at its highest value (Max), when the volume is at it highest value
+\item V(Volume(0), Height(0)): there is no height, when there is no volume
+\item V(Height(Max), Pressure(Max)): the pressure is at its highest value (Max), when the height is at it highest value
+\item V(Height(0), Pressure(0)): there is no pressure, when there is no height
+\item V(Pressure(Max), Outflow(Max)): the outflow is at its highest value (Max), when the pressure is at it highest value
+\item V(Pressure(0), Outflow(0)): there is no outflow, when there is no pressure
+\end{itemize}
+\end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{problem_extra.eps}
+\caption{Drawings of the causal model active for the extra problem}
+\end{figure}
+
 \end{document}

### 2016-10-20 search.py

@@ -2,16 +2,17 @@ from quantity import Quantity
 from problem import Problem
 from plot import Plot
 
+
 class Search:
 
-    def __init__():
-        pass
+    # def __init__(self):
+    #    pass
 
-    def equal_states(state1, state2):
-        return (len(set(state1.items()) & set(state2.items())) == len(state1.items()))
-
-    def already_in_states(state1, state2):
-        return (len(set(state1.items()) & set(state2.items())) == len(state1.items))
+    # def equal_states(state1, state2):
+    #     return (len(set(state1.items()) & set(state2.items())) == len(state1.items()))
+    #
+    # def already_in_states(state1, state2):
+    #     return (len(set(state1.items()) & set(state2.items())) == len(state1.items))
 
     # @staticmethod
     # def hash_state(state):
@@ -30,6 +31,7 @@ class Search:
             print(depth * '\t' + str(result))
             Search.recursive(problem, result, depth+1, visited_states)
 
+    @staticmethod
     def iterative(problem, state):
         result = {problem.hash_state(state): {'state': state, 'children': set([])}}
 
@@ -46,21 +48,54 @@ class Search:
 
         return result
 
-
-if __name__ == "__main__":
+def base_prob():
     inflow = Quantity("Inflow", [("Zero", False, 0), ("Plus", True, 1)])
-    tank = Quantity("Tank", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
+    volume = Quantity("Tank", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
     outflow = Quantity("Outflow", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
 
-    tank.influence(inflow, positive=True)
-    tank.influence(outflow, positive=False)
-    outflow.proportional(tank, positive=True)
+    volume.influence(inflow, positive=True)
+    volume.influence(outflow, positive=False)
+    outflow.proportional(volume, positive=True)
 
-    tank.value_constraint(outflow, "Max", "Max")
-    outflow.value_constraint(tank, "Max", "Max")
-    tank.value_constraint(outflow, "Zero", "Zero")
-    outflow.value_constraint(tank, "Zero", "Zero")
+    volume.value_constraint(outflow, "Max", "Max")
+    outflow.value_constraint(volume, "Max", "Max")
[…]

### 2016-10-20 problem.py
update log

@@ -1,26 +1,32 @@
 from copy import deepcopy, copy
-from itertools import filterfalse
+
 
 class Problem:
 
 
-    def __init__(self, quantities, fixed=False):
+    def __init__(self, quantities, fixed=False,logfile=None):
         self.quantities = quantities
         self.fixed = fixed
+        self.logfile=logfile
 
     def hash_state(self, state):
         sorted_keys = [q.name for q in self.quantities]
         result_hash = ''
-
+        
         for key in sorted_keys:
             result_hash += key + str(state[key]) + '\n'
 
         return result_hash
-
+    def state2printstring(self,state):
+        s=self.hash_state(state)
+        return s.replace("\n", ",").replace(", None","")
+        
     def succ(self, state):
         new_states = [{}]
 
         # Evolving, i.e. changing values based on derivatives
+        if self.logfile!=None:
+            print("\t Expanding the node ("+self.state2printstring(state)+") to determine its successors:")  
         for quantity in self.quantities:
             value, derivative = state[quantity.name]
             quantity_new_states = []
@@ -33,23 +39,53 @@ class Problem:
                     # print(quantity_new_states)
             new_states = quantity_new_states
             # print('NS: ' + str(new_states))
+            
+                    
+        
+        if self.logfile!=None:
+            print("\t\t for each quantity we determine the possible values in which it can trasition (using its actual value and its derivative):")    
+        if self.logfile!=None:
+            if new_states:   
+                
+                for quantity in self.quantities:
+                    value, derivative = state[quantity.name]
+                    print("\t\t\t  "+quantity.name+" : "+str(quantity.next_value(value, derivative)).replace("["," ").replace("]"," "))
+            else:
+                print("\t\t One of the quantities has no successors, thus the state is a terminal state\n")
+        
+    
 
         # Value constraints, i.e. filtering out states not matching value constraints
+        if self.logfile!=None:
+            print("\t\t For each possible combinaton of values we then check if they respect the value constraints ")
         vc_new_states = []
         for s in new_states:
             if not self.check_constraints(s):
                 vc_new_states.append(s)
         new_states = vc_new_states
-
+        if self.logfile!=None:
+            if not new_states:   
+                print("\t\t After imposing value constraints one of the quntities has no successors, thus the state is a terminal state\n")
+            else:
+                print("\t\t Now for each possible combination of values we propagate the constraints to determine the possible values the derivative can assume\n")
[…]

### 2016-10-20 search.py

@@ -30,6 +30,7 @@ class Search:
             print(depth * '\t' + str(result))
             Search.recursive(problem, result, depth+1, visited_states)
 
+    @staticmethod
     def iterative(problem, state):
         result = {problem.hash_state(state): {'state': state, 'children': set([])}}
 
@@ -61,7 +62,7 @@ if __name__ == "__main__":
     tank.value_constraint(outflow, "Zero", "Zero")
     outflow.value_constraint(tank, "Zero", "Zero")
 
-    prob1 = Problem([inflow, tank, outflow], fixed=True)
+    prob1 = Problem([inflow, tank, outflow], fixed=True,logfile=True)
 
     start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
     # start_state = {"Inflow": ["Plus", 1], "Tank": ["Plus", -1], "Outflow": ["Plus", -1]}
@@ -70,7 +71,7 @@ if __name__ == "__main__":
     # for k, v in Search.iterative(prob1, start_state).items():
     #     print(v['state'])
     print(len(Search.iterative(prob1, start_state)))
-    Plot.draw(Search.iterative(prob1, start_state), 'result.png')
+    #Plot.draw(Search.iterative(prob1, start_state), 'result.png')
     # results = prob1.succ({"Inflow": ("Zero", 1), "Tank": ("Zero", 0), "Outflow": ("Zero", 0)})
     # for res in results:
     #     print(res)

### 2016-10-21 problem.py
update log2 final changes---now we have to print it in a file

@@ -19,14 +19,14 @@ class Problem:
         return result_hash
     def state2printstring(self,state):
         s=self.hash_state(state)
-        return s.replace("\n", ",").replace(", None","")
+        return "("+s.replace("\n", " ").replace(", None","")+")"
         
     def succ(self, state):
         new_states = [{}]
 
         # Evolving, i.e. changing values based on derivatives
         if self.logfile!=None:
-            print("\t Expanding the node ("+self.state2printstring(state)+") to determine its successors:")  
+            print("--->Expanding the node ("+self.state2printstring(state)+") to determine its successors:")  
         for quantity in self.quantities:
             value, derivative = state[quantity.name]
             quantity_new_states = []
@@ -98,7 +98,8 @@ class Problem:
         for new_state in new_states:
             final_new_states += self.propagate_proportionals(state, new_state)
         
-
+        if self.logfile!=None:
+            print("---> The state"+self.state2printstring(state)+" can transition to "+str([self.state2printstring(i) for i in  final_new_states]))  
         return final_new_states
 
     def propagate_influences(self, quantity, old_state, new_state):
@@ -160,40 +161,60 @@ class Problem:
                     elif (deri_new_dict[quantity.name][0] != old_state[quantity.name][0] and interval and deri != old_state[quantity.name][1]) and self.logfile!=None:
                         print("\t\t\t\t discarding partial state "+self.state2printstring(deri_new_dict)+" because it violates continuity constraints")
             new_states = quantity_new_states
+            
         if self.logfile!=None:
             for i in new_states:
-                print("\t\t\t\t adding "+self.state2printstring(deri_new_dict)+" to the list of future states")
+                print("\t\t\t\t adding "+self.state2printstring(i)+" to the list of future states")
         return new_states
 
     def propagate_proportional(self, quantity, visited_quantities, derivatives, old_state, new_state):
-        if self.logfile!=None and quantity.proportionals:
+        if self.logfile!=None:
             print("\t\t\t\t recursive call "+quantity.name)
         visited_quantities[quantity.name] = set()
         if not quantity.hasProportionals() and not quantity.hasInfluences():
+            if self.logfile!=None :
+                print("\t\t\t\t\t the quantity receives no influence or proportionality thus we can determine its derivative arbitrarily")
             _, der = old_state[quantity.name]
             if self.fixed:
+                if self.logfile!=None :
[…]

### 2016-10-21 search.py

@@ -33,18 +33,28 @@ class Search:
     @staticmethod
     def iterative(problem, state):
         result = {problem.hash_state(state): {'state': state, 'children': set([])}}
+        if problem.logfile!= None:
+            print("The program starts to explore from the initial state "+problem.state2printstring(state)+" iteratively expanding the graph of possible beahaviours")
 
         processing_list = [state]
         while processing_list:
             processing_state = processing_list.pop(0)
+            if problem.logfile!= None:
+                print("Popping the state "+problem.state2printstring(state)+" from the list of the states to explore")
             succesors = problem.succ(processing_state)
             for succ in succesors:
                 hashed_state = problem.hash_state(succ)
                 result[problem.hash_state(processing_state)]['children'].add(hashed_state)
                 if hashed_state not in result:
+                    if problem.logfile!= None:
+                        print("Adding the state "+problem.state2printstring(state)+" to the list of the states to explore") 
                     result[hashed_state] = {'state': succ, 'children': set([])}
                     processing_list.append(succ)
-
+                elif problem.logfile!= None:
+                    print("The state "+problem.state2printstring(state)+" was already explored")
+                    
+        if problem.logfile!= None:
+                        print("The list of the states to explore is empty: terminating the program")
         return result
 
 
@@ -62,7 +72,7 @@ if __name__ == "__main__":
     tank.value_constraint(outflow, "Zero", "Zero")
     outflow.value_constraint(tank, "Zero", "Zero")
 
-    prob1 = Problem([inflow, tank, outflow], fixed=True,logfile=True)
+    prob1 = Problem([inflow, tank, outflow], fixed=False,logfile=True)
 
     start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
     # start_state = {"Inflow": ["Plus", 1], "Tank": ["Plus", -1], "Outflow": ["Plus", -1]}

### 2016-10-21 problem.py
fixed some last bugs

@@ -13,7 +13,10 @@ class Problem:
         result_hash = ''
 
         for key in sorted_keys:
-            result_hash += key + str(state[key]) + '\n'
+            result_hash += key + ": "
+            for q in state[key][:-1]:
+                result_hash += str(q) + ", "
+            result_hash += str(state[key][-1]) + '\n'
 
         return result_hash
 
@@ -41,9 +44,6 @@ class Problem:
                 vc_new_states.append(s)
         new_states = vc_new_states
 
-        # new_states = [new_states[1]]
-        # print(new_states)
-
         # Influence propogations, i.e. computing new derivatives based on influences
         for quantity in self.quantities:
             if quantity.hasInfluences():
@@ -56,6 +56,7 @@ class Problem:
                         quantity_new_states.append(value_new_dict)
                 new_states = quantity_new_states
 
+
         final_new_states = []
         for new_state in new_states:
             final_new_states += self.propagate_proportionals(state, new_state)
@@ -64,7 +65,9 @@ class Problem:
 
     def propagate_influences(self, quantity, old_state, new_state):
         new_der = set()
+        change = False
         for infl_quantity, infl_positive in quantity.influences:
+            _, der = old_state[infl_quantity.name]
             value, _ = new_state[infl_quantity.name]
             value_idx = infl_quantity.value2index(value)
             sign = infl_quantity.signs[value_idx]
@@ -72,19 +75,21 @@ class Problem:
             if sign != 0:
                 if infl_positive:
                     new_der.add(sign)
-                    # new_der.add(min(1, max(-1, old_state[quantity.name][1] + sign)))
-                else: #lif not old_state[quantity.name][1] == sign:
+                else:
                     new_der.add(-sign)
-                    # new_der.add(min(1, max(-1, old_state[quantity.name][1] - sign)))
-            #     print(new_der)
-            # else:
-            #     new_der.add(old_state[quantity.name][1])
-            #     print(new_der)
+            if (der != 0):
+                change = True
 
-        if -1 in new_der and 1 in new_der:
+        if not change:
+            new_der = set([old_state[quantity.name][1]])
+        elif -1 in new_der and 1 in new_der:
             new_der.add(0)
         elif not new_der:
             new_der.add(0)
+
+        # print(new_der)
+        # print(old_state[quantity.name][1])
+
         new_der = new_der & set([old_state[quantity.name][1], min(1, max(-1, old_state[quantity.name][1] + 1)), min(1, max(-1, old_state[quantity.name][1] - 1))])
[…]

### 2016-10-21 quantity.py

@@ -33,6 +33,14 @@ class Quantity:
     def value2index(self, value):
         return self.values.index(value)
 
+    def isMax(self, value):
+        value_idx = self.value2index(value)
+        return (value_idx == len(self.values)-1 and not self.intervals[value_idx])
+
+    def isMin(self, value):
+        value_idx = self.value2index(value)
+        return (value_idx == 0 and not self.intervals[value_idx])
+
     def next_value(self, value, derivative):
         result = []
         index = self.value2index(value)

### 2016-10-21 search.py

@@ -61,14 +61,16 @@ if __name__ == "__main__":
     tank.value_constraint(outflow, "Zero", "Zero")
     outflow.value_constraint(tank, "Zero", "Zero")
 
-    prob1 = Problem([inflow, tank, outflow], fixed=True)
+    prob1 = Problem([inflow, tank, outflow], fixed=False)
 
-    start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
-    # start_state = {"Inflow": ["Plus", 1], "Tank": ["Plus", -1], "Outflow": ["Plus", -1]}
+    start_state = {"Inflow": ["Zero", 0], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
+    start_state = {"Inflow": ["Plus", 0], "Tank": ["Zero", 1], "Outflow": ["Zero", 1]}
     # for i in prob1.succ(start_state):
     #     print(i)
     # for k, v in Search.iterative(prob1, start_state).items():
     #     print(v['state'])
+    # for i in prob1.succ(start_state):
+    #     print(i)
     print(len(Search.iterative(prob1, start_state)))
     Plot.draw(Search.iterative(prob1, start_state), 'result.png')
     # results = prob1.succ({"Inflow": ("Zero", 1), "Tank": ("Zero", 0), "Outflow": ("Zero", 0)})

### 2016-10-21 paper.tex
u

@@
 \section{State-graph}
+The algorithm from the previous section generates a list of states which are unique values of Inflow, Tank and Ouflow and their first derivatives.
+Each state has a number of reachable states after an amount of time or a change in the state from an external influence.
+For this problem the only external influence we allow is changing the derivative of the Inflow (which is opening and closing the tap).
+
+In figure \ref{state-graph} shows all possible behaviors of the system described above.
+Blue nodes are final states, which are states without other reachable states.
+Other nodes are red and each arrow represents a transition.
+
+\begin{figure}[H]\label{state-graph}
+\centering
+\includegraphics[width=\textwidth, angle=90]{result.png}
+\caption{Plot of the state-graph}
+\end{figure}
 

### 2016-10-21 problem.py
log

@@ -26,7 +26,7 @@ class Problem:
 
         # Evolving, i.e. changing values based on derivatives
         if self.logfile!=None:
-            print("--->Expanding the node ("+self.state2printstring(state)+") to determine its successors:")  
+            print("--->Expanding the node "+self.state2printstring(state)+" to determine its successors:")
         for quantity in self.quantities:
             value, derivative = state[quantity.name]
             quantity_new_states = []
@@ -43,7 +43,7 @@ class Problem:
                     
         
         if self.logfile!=None:
-            print("\t\t for each quantity we determine the possible values in which it can trasition (using its actual value and its derivative):")    
+            print("\t\t for each quantity we determine the possible values which it can trasit in (using its actual value and its derivative):")
         if self.logfile!=None:
             if new_states:   
                 

### 2016-10-21 search.py

@@ -78,8 +78,8 @@ def extra_prob():
     inflow = Quantity("Inflow", [("Zero", False, 0), ("Plus", True, 1)])
     volume = Quantity("Tank", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
     outflow = Quantity("Outflow", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
-    height = Quantity("height", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
-    pressure = Quantity("pressure", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
+    height = Quantity("Height", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
+    pressure = Quantity("Pressure", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
 
     volume.influence(inflow, positive=True)
     volume.influence(outflow, positive=False)
@@ -105,10 +105,13 @@ def extra_prob():
     return inflow, volume, outflow, height, pressure
 
 if __name__ == "__main__":
-    inflow, tank, outflow = base_prob()
-    prob1 = Problem([inflow, tank, outflow], fixed=False, logfile=True)
+    inflow, tank, outflow, height, pressure = extra_prob()
 
-    start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
+
+    #inflow, tank, outflow = base_prob()
+    prob1 = Problem([inflow, tank, outflow, height, pressure], fixed=False, logfile=True)
+
+    start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0], "Height": ["Zero", 0], "Pressure": ["Zero", 0]}
     # start_state = {"Inflow": ["Plus", 1], "Tank": ["Plus", -1], "Outflow": ["Plus", -1]}
     # for i in prob1.succ(start_state):
     #     print(i)
@@ -119,6 +122,10 @@ if __name__ == "__main__":
     # results = prob1.succ({"Inflow": ("Zero", 1), "Tank": ("Zero", 0), "Outflow": ("Zero", 0)})
     # for res in results:
     #     print(res)
+
+
+
+
 
 
 

### 2016-10-21 paper.tex
removed transition to self

@@
 \section{Algorithm}
+The goal of the algorithm is to construct a state graph based for a specific model and a given starting state.
+Constructing the state-graph is done iteratively as shown in Algorithm~\ref{alg:computing_state_graph}.
+\begin{algorithm}
+    result = \{"state": starting\_state, "children": [ ]\}
 
+    processing\_list = [starting\_state]
+
+\While{processing\_list not empty}{
+    processing\_state = pop(processing\_list)
+
+    successors = findSuccessors(processing\_state)
+
+    \For{s in succesors}{
+        \If{s not in result}{
+            add s to result
+
+            add s to processing\_list
+        }
+        add s as child to processing\_state
+    }
+}
+\caption{Computing the state-graph based on given an starting state.
+It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the succesors as children to state that is being processed, as well as adding these succesors to the processing list, but only if a successor was not already processed.}
+\label{alg:computing_state_graph}
+\end{algorithm}
+
+The core functionality of constructing the state graph is encapsulated in finding all successive states for a given state.
+The first step in finding the successive states, is propagating over all quantities in the model and computing all possible combinations of new values based on the derivatives.
+The algorithm for doing this is shown in Algorithm~\ref{alg:propogating_values}.
 \begin{algorithm}
-\If{condition}{then-block}
+new\_states = [ empty\_state ]
+
+    \For{quantity in quantities}{
+
+        p = possible next values for quantity\\
+        \ForAll{new\_states}{
+            new\_state = pop(new\_states)\\
+            \ForAll{p}{
+               add new\_state with new possible value for quantity to new\_states
+            }
+        }
+        }
+\caption{Propagate over all quantities and modify value based on the derivative.
+This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
+For each possible next value we create a new state, thus possibly creating multiple states.
+Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously}
+\label{alg:propogating_values}
+\end{algorithm}
+
+After computing all the possible successive states only with respect to the values, the algorithm performs a check whether all value constraints hold and removes the state if a state does not satisfy all value constraints.
+
+The next step in the algorithm is to determine new derivatives based on the influences.
+\begin{algorithm}
+
+\For{quantity in quantities}{
+
+    \If{quantity has influences} {
+        \ForAll{influences on quantity}{
+            new\_state = pop(new\_states)\\
+            \ForAll{p}{
+               add new\_state with new possible value for quantity to new\_states
+            }
+        }
+
+    }
+}
 \caption{}
+\label{alg:propogating_influences}
 \end{algorithm}
@@
 \centering
-\includegraphics[width=\textwidth, angle=90]{result.png}
+\includegraphics[width=\textwidth]{result.png}
 \caption{Plot of the state-graph}

### 2016-10-21 search.py

@@ -44,15 +44,16 @@ class Search:
                 print("Popping the state "+problem.state2printstring(state)+" from the list of the states to explore")
             successors = problem.succ(processing_state)
             for succ in successors:
-                hashed_state = problem.hash_state(succ)
-                result[problem.hash_state(processing_state)]['children'].add(hashed_state)
-                if hashed_state not in result:
-                    if problem.logfile!= None:
-                        print("Adding the state "+problem.state2printstring(state)+" to the list of the states to explore")
-                    result[hashed_state] = {'state': succ, 'children': set([])}
-                    processing_list.append(succ)
-                elif problem.logfile!= None:
-                    print("The state "+problem.state2printstring(state)+" was already explored")
+                if problem.hash_state(succ) != problem.hash_state(processing_state):
+                    hashed_state = problem.hash_state(succ)
+                    result[problem.hash_state(processing_state)]['children'].add(hashed_state)
+                    if hashed_state not in result:
+                        if problem.logfile!= None:
+                            print("Adding the state "+problem.state2printstring(state)+" to the list of the states to explore")
+                        result[hashed_state] = {'state': succ, 'children': set([])}
+                        processing_list.append(succ)
+                    elif problem.logfile!= None:
+                        print("The state "+problem.state2printstring(state)+" was already explored")
 
         if problem.logfile!= None:
                         print("The list of the states to explore is empty: terminating the program")
@@ -74,21 +75,17 @@ def base_prob():
 
     return inflow, volume, outflow
 
-def base_prob():
-    population = Quantity("Population", [("Zero", False, 0), ("Plus", True, 1)])
-    volume = Quantity("Tank", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
-    outflow = Quantity("Outflow", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
+def frog_prob():
+    population = Quantity("Population", [("Zero", False, 0), ("Small", True, 1), ("Medium", False, 1), ("Large", True, 1)])
+    birth = Quantity("Birth", [("Zero", False, 0), ("Plus", True, 1)])
+    death = Quantity("Death", [("Zero", False, 0), ("Plus", True, 1)])
 
[…]

### 2016-10-21 paper.tex

@@
         }
-\caption{Propagate over all quantities and modify value based on the derivative.
-This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
-For each possible next value we create a new state, thus possibly creating multiple states.
-Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously}
+\caption{Propagate over all quantities and modify value based on the derivative}
 \label{alg:propogating_values}
@@
 
-After computing all the possible successive states only with respect to the values, the algorithm performs a check whether all value constraints hold and removes the state if a state does not satisfy all value constraints.
-
-The next step in the algorithm is to determine new derivatives based on the influences.
+After changing the values according to the derivatives for each quantity in the model, we start propagating over all quantities to determine new derivatives based on the influences.
 \begin{algorithm}
-
-\For{quantity in quantities}{
-
-    \If{quantity has influences} {
-        \ForAll{influences on quantity}{
-            new\_state = pop(new\_states)\\
-            \ForAll{p}{
-               add new\_state with new possible value for quantity to new\_states
-            }
-        }
-
-    }
+\For{asd}{
+\If{condition}{then-block}
 }
 \caption{}
-\label{alg:propogating_influences}
 \end{algorithm}

### 2016-10-21 paper.tex

--- 2016-10-21 paper.tex
@@
 \caption{Computing the state-graph based on given an starting state.
-It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the succesors as children to state that is being processed, as well as adding these succesors to the processing list, but only if a successor was not already processed.}
+It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.}
 \label{alg:computing_state_graph}

### 2016-10-21 paper.tex
results and algorithm finished in paper

--- 2016-10-21 paper.tex
@@
 
-The next step in the algorithm is to determine new derivatives based on the influences.
+The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm~\ref{alg:propogating_influences}.
 \begin{algorithm}
@@
 
-    \If{quantity has influences} {
+        \ForAll{new\_states}{
+
+            new\_state = pop(new\_states)\\
+
+            possible\_derivatives = [ ]
+
             \ForAll{influences on quantity}{
-            new\_state = pop(new\_states)\\
-            \ForAll{p}{
-               add new\_state with new possible value for quantity to new\_states
+                \If{influence is not zero}{
+                    add corresponding derivative to possible\_derivative based on new\_state
                 }
+            }
+            \If{both positive and negative derivative are possible} {
+                add 0 derivative to possible\_derivative
+            }
+            \If{values of influences did not change w.r.t.\ previous state} {
+                set possible\_derivative only to old derivative
+            }
+
+            \ForAll{possible\_derivatives}{
+                add new state to new\_states with new possible derivative
             }
@@
         }
+
     }
-\caption{}
+\caption{Compute the new possible derivatives based on the influences.
+For all new states created from the previous step (value propagation), generate a list of possible derivatives.
+Then if both derivatives are possible this implies ambiguity and the 0 derivative is also a possibility.
+If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
+Finally, for each state we create multiple new states based on the possible derivatives.}
 \label{alg:propogating_influences}
@@
 
+The last step in the algorithm is computing the new derivatives based on the proportionalities.
+Again we loop over all the quantities in the model.
+Then, for each quantity we check the derivatives of the quantities it is proportional to.
+If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
+If the algorithm reaches a quantity that has not proportionalities or influences, we base its derivative on an extrogeneous influence (random).
+On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propogating the influences) and thus we do not have to go deeper into recursion.
+In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
+
+When we found possible derivatives for each quantity based on its proportionalities, we create new states for each new state that was created by the previous propogations (value, value constraints and influences).
+All the propogations steps together leave us with possible next states.
+Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
+
 \section{State-graph}
-The algorithm from the previous section generates a list of states which are unique values of Inflow, Tank and Ouflow and their first derivatives.
-Each state has a number of reachable states after an amount of time or a change in the state from an external influence.
-For this problem the only external influence we allow is changing the derivative of the Inflow (which is opening and closing the tap).
+The algorithm explained in the previous section generates a list of states, with children relations explaining the successor relations.
+Besides normal state transitions, we also have state transition due to extrogeneous influence.
+For this problem the only external influence we allow is changing the derivative of the Inflow randomly.
 
@@
 
-\begin{figure}[H]\label{state-graph}
+\begin{figure}\label{state-graph}
 \centering
-\includegraphics[width=\textwidth]{result.png}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result.png}
 \caption{Plot of the state-graph}

### 2016-10-21 search.py

@@ -90,9 +90,9 @@ def frog_prob():
 def extra_prob():
     inflow = Quantity("Inflow", [("Zero", False, 0), ("Plus", True, 1)])
     volume = Quantity("Tank", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
-    outflow = Quantity("Outflow", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
     height = Quantity("Height", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
     pressure = Quantity("Pressure", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
+    outflow = Quantity("Outflow", [("Zero", False, 0), ("Plus", True, 1), ("Max", False, 1)])
 
     volume.influence(inflow, positive=True)
     volume.influence(outflow, positive=False)
@@ -115,29 +115,24 @@ def extra_prob():
     volume.value_constraint(height, "Zero", "Zero")
     height.value_constraint(volume, "Zero", "Zero")
 
-    return inflow, volume, outflow, height, pressure
+    return inflow, volume, height, pressure, outflow
 
 if __name__ == "__main__":
-    inflow, tank, outflow, height, pressure = extra_prob()
-    prob1 = Problem([inflow, tank, height, pressure, outflow], fixed=False, logfile=True)
-    start_state = {"Inflow": ["Zero", 0], "Tank": ["Zero", 0], "Height": ["Zero", 0], "Pressure": ["Zero", 0], "Outflow": ["Zero", 0]}
+    inflow, tank, outflow = base_prob()
+    base_problem = Problem([inflow, tank, outflow], fixed=False, logfile=True)
+    base_start_state = {"Inflow": ["Zero", 0], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
+    result_base_problem = Search.iterative(base_problem, base_start_state)
+    Plot.draw(result_base_problem, 'result_base_problem.png')
 
-    population, birth, death = frog_prob()
-    prob1 = Problem([population, birth, death], fixed=True, logfile=True)
-    start_state = {"Population": ["Small", 1], "Birth": ["Plus", 1], "Death": ["Plus", 1]}
-    # start_state = {"Inflow": ["Plus", 0], "Tank": ["Zero", 1], "Outflow": ["Zero", 1]}
-    # for i in prob1.succ(start_state):
-    #     print(i)
[…]

### 2016-10-21 search.py
log2

@@ -118,12 +118,12 @@ def extra_prob():
     return inflow, volume, outflow, height, pressure
 
 if __name__ == "__main__":
-    #inflow, tank, outflow, height, pressure = extra_prob()
+    inflow, tank, outflow, height, pressure = extra_prob()
 
-    inflow, tank, outflow = base_prob()
-    prob1 = Problem([inflow, tank, outflow], fixed=False, logfile=True)
+    #inflow, tank, outflow = base_prob()
+    prob1 = Problem([inflow, tank, outflow, height, pressure], fixed=False, logfile=True)
 
-    start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0]}
+    start_state = {"Inflow": ["Zero", 1], "Tank": ["Zero", 0], "Outflow": ["Zero", 0], "Height": ["Zero", 0], "Pressure": ["Zero", 0]}
     # start_state = {"Inflow": ["Plus", 1], "Tank": ["Plus", -1], "Outflow": ["Plus", -1]}
 
     # for i in prob1.succ(start_state):

### 2016-10-21 paper.tex
log2

@@
 
-In figure \ref{state-graph} shows all possible behaviors of the system described above.
+In figure \ref{state-graph-base} shows all possible behaviors of the system described above.
 Blue nodes are final states, which are states without other reachable states.
@@
 
-\begin{figure}\label{state-graph}
+\begin{figure}\label{state-graph-base}
 \centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result.png}
-\caption{Plot of the state-graph}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result_base_problem.png}
+\caption{Plot of the state-graph of the base problem}
 \end{figure}
+
+\section{Trace}
 
@@
 
+\begin{figure}\label{state-graph-extra}
+	\centering
+	\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result_extra_problem.png}
+	\caption{Plot of the state-graph of the extra problem}
+\end{figure}
+
 \end{document}

### 2016-10-21 paper.tex
yg

--- 2016-10-21 paper.tex
@@
 
-The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm~\ref{alg:propogating_influences}.
+The next step in the algorithm is to determine new derivatives based on the influences.
 \begin{algorithm}
@@
 
-        \ForAll{new\_states}{
-
+    \If{quantity has influences} {
+        \ForAll{influences on quantity}{
             new\_state = pop(new\_states)\\
-
-            possible\_derivatives = [ ]
-
-            \ForAll{influences on quantity}{
-                \If{influence is not zero}{
-                    add corresponding derivative to possible\_derivative based on new\_state
+            \ForAll{p}{
+               add new\_state with new possible value for quantity to new\_states
             }
-            }
-            \If{both positive and negative derivative are possible} {
-                add 0 derivative to possible\_derivative
-            }
-            \If{values of influences did not change w.r.t.\ previous state} {
-                set possible\_derivative only to old derivative
-            }
-
-            \ForAll{possible\_derivatives}{
-                add new state to new\_states with new possible derivative
         }
@@
     }
-
 }
-\caption{Compute the new possible derivatives based on the influences.
-For all new states created from the previous step (value propagation), generate a list of possible derivatives.
-Then if both derivatives are possible this implies ambiguity and the 0 derivative is also a possibility.
-If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
-Finally, for each state we create multiple new states based on the possible derivatives.}
+\caption{}
 \label{alg:propogating_influences}
 \end{algorithm}
-
-The last step in the algorithm is computing the new derivatives based on the proportionalities.
-Again we loop over all the quantities in the model.
-Then, for each quantity we check the derivatives of the quantities it is proportional to.
-If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
-If the algorithm reaches a quantity that has not proportionalities or influences, we base its derivative on an extrogeneous influence (random).
-On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propogating the influences) and thus we do not have to go deeper into recursion.
-In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
-
-When we found possible derivatives for each quantity based on its proportionalities, we create new states for each new state that was created by the previous propogations (value, value constraints and influences).
-All the propogations steps together leave us with possible next states.
-Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
+\section{Modelling choices and possible extensions}
+The algorithm implemented is very general and allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.\newline
+First of all we chose not to implement all the inequality constraints.
+The main reason for not implementing this, is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
+Therefore we decided to implement only the most important ones, Such as the continuity constraints.\newline
+We then decided not to use higher order derivatives.
+However because of the generality of the algorithm, we cold easily have implemented them adding for each quantity $Q$ an another quantity $\partial Q$ and imposing constraints.
+\newline 
+Finally we only implemented random and static derivative for the inflow, but we could have easily implemented others exogenous behaviours.
 
 \section{State-graph}
-The algorithm explained in the previous section generates a list of states, with children relations explaining the successor relations.
-Besides normal state transitions, we also have state transition due to extrogeneous influence.
-For this problem the only external influence we allow is changing the derivative of the Inflow randomly.
+The algorithm from the previous section generates a list of states which are unique values of Inflow, Tank and Ouflow and their first derivatives.
+Each state has a number of reachable states after an amount of time or a change in the state from an external influence.
+For this problem the only external influence we allow is changing the derivative of the Inflow (which is opening and closing the tap).
 
@@
 
-\begin{figure}\label{state-graph}
+\begin{figure}[H]\label{state-graph}
 \centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result.png}
+\includegraphics[width=\textwidth]{result.png}
 \caption{Plot of the state-graph}

### 2016-10-21 paper.tex

--- 2016-10-21 paper.tex
@@
 \end{algorithm}
+\section{Modelling choices and possible extensions}
+The algorithm implemented is very general and allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.\newline
+First of all we chose not to implement all the inequality constraints.
+The main reason for not implementing this, is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
+Therefore we decided to implement only the most important ones, Such as the continuity constraints.\newline
+We then decided not to use higher order derivatives.
+However because of the generality of the algorithm, we cold easily have implemented them adding for each quantity $Q$ an another quantity $\partial Q$ and imposing constraints.
+\newline 
+Finally we only implemented random and static derivative for the inflow, but we could have easily implemented others exogenous behaviours.
 

### 2016-10-21 paper.tex
paper

@@
 Constructing the state-graph is done iteratively as shown in Algorithm~\ref{alg:computing_state_graph}.
-\begin{algorithm}
+
+\vspace{12pt}
+\begin{algorithm}[H]
     result = \{"state": starting\_state, "children": [ ]\}
@@
 }
-\caption{Computing the state-graph based on given an starting state.
-It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.}
+\caption{Search in the state-graph}
 \label{alg:computing_state_graph}
 \end{algorithm}
+\vspace{12pt}
+
+Algorithm \ref{alg:computing_state_graph} computing the state-graph based on given an starting state.
+It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.
 
@@
 The first step in finding the successive states, is propagating over all quantities in the model and computing all possible combinations of new values based on the derivatives.
-The algorithm for doing this is shown in Algorithm~\ref{alg:propogating_values}.
-\begin{algorithm}
+The algorithm for doing this is shown in Algorithm \ref{alg:propogating_values}.
+
+\vspace{12pt}
+\begin{algorithm}[H]
 new\_states = [ empty\_state ]
@@
         }
-\caption{Propagate over all quantities and modify value based on the derivative.
+\caption{Values propagation}
+\label{alg:propogating_values}
+\end{algorithm}
+\vspace{12pt}
+
+Algorithm \ref{alg:propogating_values} propagate over all quantities and modify value based on the derivative.
 This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
 For each possible next value we create a new state, thus possibly creating multiple states.
-Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously}
-\label{alg:propogating_values}
-\end{algorithm}
+Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously.
 
@@
 
-The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm~\ref{alg:propogating_influences}.
-\begin{algorithm}
+The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm \ref{alg:propogating_influences}.
+
+\vspace{12pt}
+\begin{algorithm}[H]
 
@@
     }
-\caption{Compute the new possible derivatives based on the influences.
+\caption{Influences propagation}
+\label{alg:propogating_influences}
+\end{algorithm}
+\vspace{12pt}
+
+Algorithm \ref{alg:propogating_influences} computes the new possible derivatives based on the influences.
 For all new states created from the previous step (value propagation), generate a list of possible derivatives.
@@
 If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
-Finally, for each state we create multiple new states based on the possible derivatives.}
-\label{alg:propogating_influences}
+Finally, for each state we create multiple new states based on the possible derivatives.
+
+\vspace{12pt}
+\begin{algorithm}[H]
+	TODO
+	\caption{Proportionals propagation}
+	\label{alg:propogating_proportionals}
 \end{algorithm}
-\section{Modelling choices and possible extensions}
-The algorithm implemented is very general and allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.\newline
-First of all we chose not to implement all the inequality constraints.
-The main reason for not implementing this, is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
-Therefore we decided to implement only the most important ones, Such as the continuity constraints.\newline
-We then decided not to use higher order derivatives.
-However because of the generality of the algorithm, we cold easily have implemented them adding for each quantity $Q$ an another quantity $\partial Q$ and imposing constraints.
-\newline 
-Finally we only implemented random and static derivative for the inflow, but we could have easily implemented others exogenous behaviours.
+\vspace{12pt}
 
-The last step in the algorithm is computing the new derivatives based on the proportionalities.
+The last step in the algorithm is computing the new derivatives based on the proportionates.
 Again we loop over all the quantities in the model.
@@
 If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
-If the algorithm reaches a quantity that has not proportionalities or influences, we base its derivative on an extrogeneous influence (random).
-On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propogating the influences) and thus we do not have to go deeper into recursion.
+If the algorithm reaches a quantity that has not proportionates or influences, we base its derivative on an extrogeneous influence (random).
+On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propagating the influences) and thus we do not have to go deeper into recursion.
 In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
 
-When we found possible derivatives for each quantity based on its proportionalities, we create new states for each new state that was created by the previous propogations (value, value constraints and influences).
-All the propogations steps together leave us with possible next states.
+When we found possible derivatives for each quantity based on its proportionates, we create new states for each new state that was created by the previous propagations (value, value constraints and influences).
+All the propagations steps together leave us with possible next states.
 Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
+
+\subsection{Modelling choices and possible extensions}
+The algorithm implemented is very general and allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.
+
+First of all we chose not to implement all the inequality constraints.
+The main reason for not implementing this, is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
+Therefore we decided to implement only the most important ones, Such as the continuity constraints.
+
+We then decided not to use higher order derivatives.
+However because of the generality of the algorithm, we cold easily have implemented them adding for each quantity $Q$ an another quantity $\partial Q$ and imposing constraints.
+Finally we only implemented random and static derivative for the inflow, but we could have easily implemented others exogenous behaviors.
 
@@
 
-In figure \ref{state-graph-base} shows all possible behaviors of the system described above.
-Blue nodes are final states, which are states without other reachable states.
-Other nodes are red and each arrow represents a transition.
+In figure \ref{state-graph-base} shows all possible behaviors of the system described above. Nodes are states and each arrow represents a transition to every reachable states.
 

### 2016-10-21 paper.tex
paper

@@
 
+\newpage
 \section{Algorithm}
@@
 
-In figure \ref{state-graph-base} shows all possible behaviors of the system described above. Nodes are states and each arrow represents a transition to every reachable states.
+In figure \ref{state-graph-base} shows all possible behaviors of the system described above.
+Nodes are states and each arrow represents a transition to every reachable states.
+Nodes are labeled with the description of the current value of each quantity and its derivative.
 
-\begin{figure}\label{state-graph-base}
+\begin{figure}[H]\label{state-graph-base}
 \centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result_base_problem.png}
-\caption{Plot of the state-graph of the base problem}
+\includegraphics[width=\textwidth,height=7.7in,keepaspectratio]{result_base_problem.png}
+\caption{Plot of the state-graph for the base problem}
 \end{figure}
@@
 \item \textbf{Dependencies:}
-
 \begin{itemize}
@@
 \centering
-\includegraphics[scale=0.5]{problem_extra.eps}
+\includegraphics[scale=0.4]{problem_extra.eps}
 \caption{Drawings of the causal model active for the extra problem}
@@
 
-\begin{figure}\label{state-graph-extra}
+\begin{figure}[H]\label{state-graph-extra}
 	\centering
-	\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result_extra_problem.png}
-	\caption{Plot of the state-graph of the extra problem}
+	\includegraphics[width=\textwidth,height=9.5in,keepaspectratio]{result_extra_problem.png}
+	\caption{Plot of the state-graph for the extra problem}
 \end{figure}

### 2016-10-21 paper.tex
trolololo
paper

@@
 The goal of the algorithm is to construct a state graph based for a specific model and a given starting state.
+We decided only to model the states that are reachable from the starting state.
+The reason for doing so is that we only want physically plausible states.
+That is we don't want disconnected states or states that cannot be reached by the system.
+
 Constructing the state-graph is done iteratively as shown in Algorithm~\ref{alg:computing_state_graph}.
@@
 
-Algorithm \ref{alg:computing_state_graph} computing the state-graph based on given an starting state.
+Algorithm \ref{alg:computing_state_graph} computes the state-graph based on given an starting state.
 It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.
@@
 
-Algorithm \ref{alg:propogating_values} propagate over all quantities and modify value based on the derivative.
+Algorithm \ref{alg:propogating_values} propagates over all quantities and modify value based on the derivative.
 This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
@@
 If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
-If the algorithm reaches a quantity that has not proportionates or influences, we base its derivative on an extrogeneous influence (random).
+If the algorithm reaches a quantity that has not proportionates or influences, we base its derivative on an exogenous influence (random).
 On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propagating the influences) and thus we do not have to go deeper into recursion.
@@
 \subsection{Modelling choices and possible extensions}
-The algorithm implemented is very general and allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.
+The implemented algorithm is very general and it allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.
 
 First of all we chose not to implement all the inequality constraints.
-The main reason for not implementing this, is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
+The main reason for not including these is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
 Therefore we decided to implement only the most important ones, Such as the continuity constraints.
@@
 We then decided not to use higher order derivatives.
-However because of the generality of the algorithm, we cold easily have implemented them adding for each quantity $Q$ an another quantity $\partial Q$ and imposing constraints.
+However because of the generality of the algorithm, we could easily have implemented them adding, for each quantity $Q$, an another quantity $\partial Q$, and imposing constraints.
 Finally we only implemented random and static derivative for the inflow, but we could have easily implemented others exogenous behaviors.
@@
 The algorithm explained in the previous section generates a list of states, with children relations explaining the successor relations.
-Besides normal state transitions, we also have state transition due to extrogeneous influence.
+Besides normal state transitions, we also have state transition due to exogenous influence.
 For this problem the only external influence we allow is changing the derivative of the Inflow randomly.
@@
 In figure \ref{state-graph-base} shows all possible behaviors of the system described above.
-Nodes are states and each arrow represents a transition to every reachable states.
+Nodes are states and arrows represent transitions from a state to every reachable states starting from that source.
 Nodes are labeled with the description of the current value of each quantity and its derivative.

### 2016-10-21 paper.tex
trolololo

--- 2016-10-21 paper.tex
@@
 
+The trace reports how the algorithm expands the state graph and how it determines all the possible transitions for a state.
+
+In particular the trace gives insights on the procedure for computing the possible successors.
+It shows how the values are assigned and how the constraints are propagated.
+It also tells when we discard a partial assignment because it violates value or continuity constraints.
+
+However we decided to show the big picture of what the algorithm is doing than to print every single step in order keep the file readable.
+Thus we don't report all the intermediate steps and computations the programs is performing highlighting only the principal steps.
+
 \section{Extra Problem}

### 2016-10-21 paper.tex
paper

--- 2016-10-21 paper.tex
@@
 \item P+(Volume, Outflow): outflow changes are proportional to volume changes
-\item V(Volume(Max), Outflow(Max)): the outflow is at its highest value (Max), when the volume is at it highest value
-\item V(Volume(0), Outflow(0)): there is no outflow, when there is no volume
+\item VC(Volume(Max), Outflow(Max)): the outflow is at its highest value (Max), when the volume is at it highest value
+\item VC(Volume(0), Outflow(0)): there is no outflow, when there is no volume
 \end{itemize}
@@
 
-\subsection{Modelling choices and possible extensions}
+\subsection{Modeling choices and possible extensions}
 The implemented algorithm is very general and it allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.
@@
 \section{Trace}
+In order to debug the code and let users to better understand what the algorithm do we print a trace.
+The trace of the base problem is in \texttt{base\_trace} file and the trace of the extra problem is in \texttt{extra\_trace} file.
 
@@
 \item P+(Pressure, Outflow): outflow changes are proportional to pressure changes
-\item V(Volume(Max), Height(Max)): the height is at its highest value (Max), when the volume is at it highest value
-\item V(Volume(0), Height(0)): there is no height, when there is no volume
-\item V(Height(Max), Pressure(Max)): the pressure is at its highest value (Max), when the height is at it highest value
-\item V(Height(0), Pressure(0)): there is no pressure, when there is no height
-\item V(Pressure(Max), Outflow(Max)): the outflow is at its highest value (Max), when the pressure is at it highest value
-\item V(Pressure(0), Outflow(0)): there is no outflow, when there is no pressure
+\item VC(Volume(Max), Height(Max)): the height is at its highest value (Max), when the volume is at it highest value
+\item VC(Volume(0), Height(0)): there is no height, when there is no volume
+\item VC(Height(Max), Pressure(Max)): the pressure is at its highest value (Max), when the height is at it highest value
+\item VC(Height(0), Pressure(0)): there is no pressure, when there is no height
+\item VC(Pressure(Max), Outflow(Max)): the outflow is at its highest value (Max), when the pressure is at it highest value
+\item VC(Pressure(0), Outflow(0)): there is no outflow, when there is no pressure
 \end{itemize}
@@
 \centering
-\includegraphics[scale=0.4]{problem_extra.eps}
+\includegraphics[scale=0.5]{problem_extra.eps}
 \caption{Drawings of the causal model active for the extra problem}

### 2016-10-21 paper.tex
trolololo
paper

--- 2016-10-21 paper.tex
@@
 
-\newpage
 \section{Algorithm}
 The goal of the algorithm is to construct a state graph based for a specific model and a given starting state.
-We decided only to model the states that are reachable from the starting state.
-The reason for doing so is that we only want physically plausible states.
-That is we don't want disconnected states or states that cannot be reached by the system.
-
 Constructing the state-graph is done iteratively as shown in Algorithm~\ref{alg:computing_state_graph}.
-
-\vspace{12pt}
-\begin{algorithm}[H]
+\begin{algorithm}
     result = \{"state": starting\_state, "children": [ ]\}
@@
 }
-\caption{Search in the state-graph}
+\caption{Computing the state-graph based on given an starting state.
+It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the succesors as children to state that is being processed, as well as adding these succesors to the processing list, but only if a successor was not already processed.}
 \label{alg:computing_state_graph}
 \end{algorithm}
-\vspace{12pt}
-
-Algorithm \ref{alg:computing_state_graph} computes the state-graph based on given an starting state.
-It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.
 
@@
 The first step in finding the successive states, is propagating over all quantities in the model and computing all possible combinations of new values based on the derivatives.
-The algorithm for doing this is shown in Algorithm \ref{alg:propogating_values}.
-
-\vspace{12pt}
-\begin{algorithm}[H]
+The algorithm for doing this is shown in Algorithm~\ref{alg:propogating_values}.
+\begin{algorithm}
 new\_states = [ empty\_state ]
@@
         }
-\caption{Values propagation}
+\caption{Propagate over all quantities and modify value based on the derivative.
+This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
+For each possible next value we create a new state, thus possibly creating multiple states.
+Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously}
 \label{alg:propogating_values}
 \end{algorithm}
-\vspace{12pt}
-
-Algorithm \ref{alg:propogating_values} propagates over all quantities and modify value based on the derivative.
-This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
-For each possible next value we create a new state, thus possibly creating multiple states.
-Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously.
 
@@
 
-The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm \ref{alg:propogating_influences}.
-
-\vspace{12pt}
-\begin{algorithm}[H]
+The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm~\ref{alg:propogating_influences}.
+\begin{algorithm}
 
@@
     }
-\caption{Influences propagation}
-\label{alg:propogating_influences}
-\end{algorithm}
-\vspace{12pt}
-
-Algorithm \ref{alg:propogating_influences} computes the new possible derivatives based on the influences.
+\caption{Compute the new possible derivatives based on the influences.
 For all new states created from the previous step (value propagation), generate a list of possible derivatives.
@@
 If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
-Finally, for each state we create multiple new states based on the possible derivatives.
+Finally, for each state we create multiple new states based on the possible derivatives.}
+\label{alg:propogating_influences}
+\end{algorithm}
 
-\vspace{12pt}
-\begin{algorithm}[H]
-	TODO
-	\caption{Proportionals propagation}
-	\label{alg:propogating_proportionals}
-\end{algorithm}
-\vspace{12pt}
-
-The last step in the algorithm is computing the new derivatives based on the proportionates.
+The last step in the algorithm is computing the new derivatives based on the proportionalities.
 Again we loop over all the quantities in the model.
@@
 If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
-If the algorithm reaches a quantity that has not proportionates or influences, we base its derivative on an exogenous influence (random).
-On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propagating the influences) and thus we do not have to go deeper into recursion.
+If the algorithm reaches a quantity that has not proportionalities or influences, we base its derivative on an extrogeneous influence (random).
+On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propogating the influences) and thus we do not have to go deeper into recursion.
 In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
 
-When we found possible derivatives for each quantity based on its proportionates, we create new states for each new state that was created by the previous propagations (value, value constraints and influences).
-All the propagations steps together leave us with possible next states.
+When we found possible derivatives for each quantity based on its proportionalities, we create new states for each new state that was created by the previous propogations (value, value constraints and influences).
+All the propogations steps together leave us with possible next states.
 Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
-
-\subsection{Modelling choices and possible extensions}
-The implemented algorithm is very general and it allows to model an arbitrarily complex model. Despite that we decided not to include certain features in our program.
-
-First of all we chose not to implement all the inequality constraints.
-The main reason for not including these is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
-Therefore we decided to implement only the most important ones, Such as the continuity constraints.
-
-We then decided not to use higher order derivatives.
-However because of the generality of the algorithm, we could easily have implemented them adding, for each quantity $Q$, an another quantity $\partial Q$, and imposing constraints.
-Finally we only implemented random and static derivative for the inflow, but we could have easily implemented others exogenous behaviors.
 
@@
 The algorithm explained in the previous section generates a list of states, with children relations explaining the successor relations.
-Besides normal state transitions, we also have state transition due to exogenous influence.
+Besides normal state transitions, we also have state transition due to extrogeneous influence.
 For this problem the only external influence we allow is changing the derivative of the Inflow randomly.
 
-In figure \ref{state-graph-base} shows all possible behaviors of the system described above.
-Nodes are states and arrows represent transitions from a state to every reachable states starting from that source.
-Nodes are labeled with the description of the current value of each quantity and its derivative.
+In figure \ref{state-graph} shows all possible behaviors of the system described above.
+Blue nodes are final states, which are states without other reachable states.
+Other nodes are red and each arrow represents a transition.
 
-\begin{figure}[H]\label{state-graph-base}
+\begin{figure}\label{state-graph}
 \centering
-\includegraphics[width=\textwidth,height=7.7in,keepaspectratio]{result_base_problem.png}
-\caption{Plot of the state-graph for the base problem}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result.png}
+\caption{Plot of the state-graph}
 \end{figure}
-
-\section{Trace}
-
-The trace reports how the algorithm expands the state graph and how it determines all the possible transitions for a state.
-
-In particular the trace gives insights on the procedure for computing the possible successors.
-It shows how the values are assigned and how the constraints are propagated.
-It also tells when we discard a partial assignment because it violates value or continuity constraints.
-
-However we decided to show the big picture of what the algorithm is doing than to print every single step in order keep the file readable.
-Thus we don't report all the intermediate steps and computations the programs is performing highlighting only the principal steps.
 
@@
 \item \textbf{Dependencies:}
+
 \begin{itemize}
@@
 
-\begin{figure}[H]\label{state-graph-extra}
+\end{document}
+
+\documentclass[a4paper]{article}
+\usepackage[margin=1in]{geometry}
+\usepackage[english]{babel}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath,amsthm,amssymb}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{pdfpages}
+\usepackage{float, algorithmic, algorithm2e, program}
+\usepackage{tabularx}
+\usepackage{longtable}
+\usepackage{epstopdf}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}[section]
+
+\title{Bathtub Qualitative Reasoning}
+\author{Nicola De Cao, Luca Falorsi, Govert Verkes}
+\date{\today}
+
+\begin{document}
+\maketitle
+
+\section{Base Problem}
+
+\begin{itemize}
+\item \textbf{Quantities:}
+\begin{itemize}
+\item Inflow (of water into the container) $\in [0, +]$
+\item Outflow (of water out of the container) $\in [0, +, Max]$
+\item Volume (of the water in the container) $\in [0, +, Max]$
+\end{itemize}
+
+\item \textbf{Dependencies:}
+
+\begin{itemize}
+\item I+(Inflow, Volume): the amount of inflow increases the volume
+\item I-(Outflow, Volume): the amount of outflow decreases the volume
+\item P+(Volume, Outflow): outflow changes are proportional to volume changes
+\item V(Volume(Max), Outflow(Max)): the outflow is at its highest value (Max), when the volume is at it highest value
+\item V(Volume(0), Outflow(0)): there is no outflow, when there is no volume
+\end{itemize}
+\end{itemize}
+
+\begin{figure}[H]
 \centering
-	\includegraphics[width=\textwidth,height=9.5in,keepaspectratio]{result_extra_problem.png}
-	\caption{Plot of the state-graph for the extra problem}
+\includegraphics[scale=0.5]{problem1.eps}
+\caption{Drawings of the causal model active for the problem}
+\end{figure}
+
+\section{Algorithm}
+The goal of the algorithm is to construct a state graph based for a specific model and a given starting state.
+Constructing the state-graph is done iteratively as shown in Algorithm~\ref{alg:computing_state_graph}.
+\begin{algorithm}
+    result = \{"state": starting\_state, "children": [ ]\}
+
+    processing\_list = [starting\_state]
+
+\While{processing\_list not empty}{
+    processing\_state = pop(processing\_list)
+
+    successors = findSuccessors(processing\_state)
+
+    \For{s in succesors}{
+        \If{s not in result}{
+            add s to result
+
+            add s to processing\_list
+        }
+        add s as child to processing\_state
+    }
+}
+\caption{Computing the state-graph based on given an starting state.
+It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.}
+\label{alg:computing_state_graph}
+\end{algorithm}
+
+The core functionality of constructing the state graph is encapsulated in finding all successive states for a given state.
+The first step in finding the successive states, is propagating over all quantities in the model and computing all possible combinations of new values based on the derivatives.
+The algorithm for doing this is shown in Algorithm~\ref{alg:propogating_values}.
+\begin{algorithm}
+new\_states = [ empty\_state ]
+
+    \For{quantity in quantities}{
+
+        p = possible next values for quantity\\
+        \ForAll{new\_states}{
+            new\_state = pop(new\_states)\\
+            \ForAll{p}{
+               add new\_state with new possible value for quantity to new\_states
+            }
+        }
+        }
+\caption{Propagate over all quantities and modify value based on the derivative.
+This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
+For each possible next value we create a new state, thus possibly creating multiple states.
+Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously}
+\label{alg:propogating_values}
+\end{algorithm}
+
+After computing all the possible successive states only with respect to the values, the algorithm performs a check whether all value constraints hold and removes the state if a state does not satisfy all value constraints.
+
+The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm~\ref{alg:propogating_influences}.
+\begin{algorithm}
+
+    \For{quantity in quantities}{
+
+        \ForAll{new\_states}{
+
+            new\_state = pop(new\_states)\\
+
+            possible\_derivatives = [ ]
+
+            \ForAll{influences on quantity}{
+                \If{influence is not zero}{
+                    add corresponding derivative to possible\_derivative based on new\_state
+                }
+            }
+            \If{both positive and negative derivative are possible} {
+                add 0 derivative to possible\_derivative
+            }
+            \If{values of influences did not change w.r.t.\ previous state} {
+                set possible\_derivative only to old derivative
+            }
+
+            \ForAll{possible\_derivatives}{
+                add new state to new\_states with new possible derivative
+            }
+
+        }
+
+    }
+\caption{Compute the new possible derivatives based on the influences.
+For all new states created from the previous step (value propagation), generate a list of possible derivatives.
+Then if both derivatives are possible this implies ambiguity and the 0 derivative is also a possibility.
+If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
+Finally, for each state we create multiple new states based on the possible derivatives.}
+\label{alg:propogating_influences}
+\end{algorithm}
+
+The last step in the algorithm is computing the new derivatives based on the proportionalities.
+Again we loop over all the quantities in the model.
+Then, for each quantity we check the derivatives of the quantities it is proportional to.
+If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
+If the algorithm reaches a quantity that has not proportionalities or influences, we base its derivative on an extrogeneous influence (random).
+On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propogating the influences) and thus we do not have to go deeper into recursion.
+In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
+
+When we found possible derivatives for each quantity based on its proportionalities, we create new states for each new state that was created by the previous propogations (value, value constraints and influences).
+All the propogations steps together leave us with possible next states.
+Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
+
+\section{State-graph}
+The algorithm explained in the previous section generates a list of states, with children relations explaining the successor relations.
+Besides normal state transitions, we also have state transition due to extrogeneous influence.
+For this problem the only external influence we allow is changing the derivative of the Inflow randomly.
+
+In figure \ref{state-graph} shows all possible behaviors of the system described above.
+Blue nodes are final states, which are states without other reachable states.
+Other nodes are red and each arrow represents a transition.
+
+\begin{figure}\label{state-graph}
+\centering
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result.png}
+\caption{Plot of the state-graph}
+\end{figure}
+
+\section{Extra Problem}
+
+\begin{itemize}
+\item \textbf{Quantities:}
+\begin{itemize}
+\item Inflow (of water into the container) $\in [0, +]$
+\item Outflow (of water out of the container) $\in [0, +, Max]$
+\item Volume (of the water in the container) $\in [0, +, Max]$
+\item Height (of the water column in the container) $\in [0, +, Max]$
+\item Pressure (of the water column at the bottom of the container) $\in [0, +, Max]$
+\end{itemize}
+
+\item \textbf{Dependencies:}
+
+\begin{itemize}
+\item I+(Inflow, Volume): the amount of inflow increases the volume
+\item I-(Outflow, Volume): the amount of outflow decreases the volume
+\item P+(Volume, Height): height changes are proportional to volume changes
+\item P+(Height, Pressure): pressure changes are proportional to height changes
+\item P+(Pressure, Outflow): outflow changes are proportional to pressure changes
+\item V(Volume(Max), Height(Max)): the height is at its highest value (Max), when the volume is at it highest value
+\item V(Volume(0), Height(0)): there is no height, when there is no volume
+\item V(Height(Max), Pressure(Max)): the pressure is at its highest value (Max), when the height is at it highest value
+\item V(Height(0), Pressure(0)): there is no pressure, when there is no height
+\item V(Pressure(Max), Outflow(Max)): the outflow is at its highest value (Max), when the pressure is at it highest value
+\item V(Pressure(0), Outflow(0)): there is no outflow, when there is no pressure
+\end{itemize}
+\end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{problem_extra.eps}
+\caption{Drawings of the causal model active for the extra problem}
 \end{figure}
@@
 \end{document}
+
+\documentclass[a4paper]{article}
+\usepackage[margin=1in]{geometry}
+\usepackage[english]{babel}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath,amsthm,amssymb}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{pdfpages}
+\usepackage{float, algorithmic, algorithm2e, program}
+\usepackage{tabularx}
+\usepackage{longtable}
+\usepackage{epstopdf}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}[section]
+
+\title{Bathtub Qualitative Reasoning}
+\author{Nicola De Cao, Luca Falorsi, Govert Verkes}
+\date{\today}
+
+\begin{document}
+\maketitle
+
+\section{Base Problem}
+
+\begin{itemize}
+\item \textbf{Quantities:}
+\begin{itemize}
+\item Inflow (of water into the container) $\in [0, +]$
+\item Outflow (of water out of the container) $\in [0, +, Max]$
+\item Volume (of the water in the container) $\in [0, +, Max]$
+\end{itemize}
+
+\item \textbf{Dependencies:}
+
+\begin{itemize}
+\item I+(Inflow, Volume): the amount of inflow increases the volume
+\item I-(Outflow, Volume): the amount of outflow decreases the volume
+\item P+(Volume, Outflow): outflow changes are proportional to volume changes
+\item V(Volume(Max), Outflow(Max)): the outflow is at its highest value (Max), when the volume is at it highest value
+\item V(Volume(0), Outflow(0)): there is no outflow, when there is no volume
+\end{itemize}
+\end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{problem1.eps}
+\caption{Drawings of the causal model active for the problem}
+\end{figure}
+
+\section{Algorithm}
+The goal of the algorithm is to construct a state graph based for a specific model and a given starting state.
+Constructing the state-graph is done iteratively as shown in Algorithm~\ref{alg:computing_state_graph}.
+\begin{algorithm}
+    result = \{"state": starting\_state, "children": [ ]\}
+
+    processing\_list = [starting\_state]
+
+\While{processing\_list not empty}{
+    processing\_state = pop(processing\_list)
+
+    successors = findSuccessors(processing\_state)
+
+    \For{s in succesors}{
+        \If{s not in result}{
+            add s to result
+
+            add s to processing\_list
+        }
+        add s as child to processing\_state
+    }
+}
+\caption{Computing the state-graph based on given an starting state.
+It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.}
+\label{alg:computing_state_graph}
+\end{algorithm}
+
+The core functionality of constructing the state graph is encapsulated in finding all successive states for a given state.
+The first step in finding the successive states, is propagating over all quantities in the model and computing all possible combinations of new values based on the derivatives.
+The algorithm for doing this is shown in Algorithm~\ref{alg:propogating_values}.
+\begin{algorithm}
+new\_states = [ empty\_state ]
+
+    \For{quantity in quantities}{
+
+        p = possible next values for quantity\\
+        \ForAll{new\_states}{
+            new\_state = pop(new\_states)\\
+            \ForAll{p}{
+               add new\_state with new possible value for quantity to new\_states
+            }
+        }
+        }
+\caption{Propagate over all quantities and modify value based on the derivative.
+This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
+For each possible next value we create a new state, thus possibly creating multiple states.
+Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously}
+\label{alg:propogating_values}
+\end{algorithm}
+
+After computing all the possible successive states only with respect to the values, the algorithm performs a check whether all value constraints hold and removes the state if a state does not satisfy all value constraints.
+
+The next step in the algorithm is to determine new derivatives based on the influences, the algorithm for doing this is shown in Algorithm~\ref{alg:propogating_influences}.
+\begin{algorithm}
+
+    \For{quantity in quantities}{
+
+        \ForAll{new\_states}{
+
+            new\_state = pop(new\_states)\\
+
+            possible\_derivatives = [ ]
+
+            \ForAll{influences on quantity}{
+                \If{influence is not zero}{
+                    add corresponding derivative to possible\_derivative based on new\_state
+                }
+            }
+            \If{both positive and negative derivative are possible} {
+                add 0 derivative to possible\_derivative
+            }
+            \If{values of influences did not change w.r.t.\ previous state} {
+                set possible\_derivative only to old derivative
+            }
+
+            \ForAll{possible\_derivatives}{
+                add new state to new\_states with new possible derivative
+            }
+
+        }
+
+    }
+\caption{Compute the new possible derivatives based on the influences.
+For all new states created from the previous step (value propagation), generate a list of possible derivatives.
+Then if both derivatives are possible this implies ambiguity and the 0 derivative is also a possibility.
+If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
+Finally, for each state we create multiple new states based on the possible derivatives.}
+\label{alg:propogating_influences}
+\end{algorithm}
+
+The last step in the algorithm is computing the new derivatives based on the proportionalities.
+Again we loop over all the quantities in the model.
+Then, for each quantity we check the derivatives of the quantities it is proportional to.
+If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
+If the algorithm reaches a quantity that has not proportionalities or influences, we base its derivative on an extrogeneous influence (random).
+On the other hand, if the algorithm reaches a quantity that only has influences, then we know that the derivative of this quantity is already determined in the previous step (propogating the influences) and thus we do not have to go deeper into recursion.
+In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
+
+When we found possible derivatives for each quantity based on its proportionalities, we create new states for each new state that was created by the previous propogations (value, value constraints and influences).
+All the propogations steps together leave us with possible next states.
+Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
+
+\section{State-graph}
+The algorithm explained in the previous section generates a list of states, with children relations explaining the successor relations.
+Besides normal state transitions, we also have state transition due to extrogeneous influence.
+For this problem the only external influence we allow is changing the derivative of the Inflow randomly.
+
+In figure \ref{state-graph} shows all possible behaviors of the system described above.
+Blue nodes are final states, which are states without other reachable states.
+Other nodes are red and each arrow represents a transition.
+
+\begin{figure}\label{state-graph}
+\centering
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{result.png}
+\caption{Plot of the state-graph}
+\end{figure}
+
+\section{Extra Problem}
+
+\begin{itemize}
+\item \textbf{Quantities:}
+\begin{itemize}
+\item Inflow (of water into the container) $\in [0, +]$
+\item Outflow (of water out of the container) $\in [0, +, Max]$
+\item Volume (of the water in the container) $\in [0, +, Max]$
+\item Height (of the water column in the container) $\in [0, +, Max]$
+\item Pressure (of the water column at the bottom of the container) $\in [0, +, Max]$
+\end{itemize}
+
+\item \textbf{Dependencies:}
+
+\begin{itemize}
+\item I+(Inflow, Volume): the amount of inflow increases the volume
+\item I-(Outflow, Volume): the amount of outflow decreases the volume
+\item P+(Volume, Height): height changes are proportional to volume changes
+\item P+(Height, Pressure): pressure changes are proportional to height changes
+\item P+(Pressure, Outflow): outflow changes are proportional to pressure changes
+\item V(Volume(Max), Height(Max)): the height is at its highest value (Max), when the volume is at it highest value
+\item V(Volume(0), Height(0)): there is no height, when there is no volume
+\item V(Height(Max), Pressure(Max)): the pressure is at its highest value (Max), when the height is at it highest value
+\item V(Height(0), Pressure(0)): there is no pressure, when there is no height
+\item V(Pressure(Max), Outflow(Max)): the outflow is at its highest value (Max), when the pressure is at it highest value
+\item V(Pressure(0), Outflow(0)): there is no outflow, when there is no pressure
+\end{itemize}
+\end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{problem_extra.eps}
+\caption{Drawings of the causal model active for the extra problem}
+\end{figure}
+
+\end{document}

### 2016-10-21 paper.tex

--- 2016-10-21 paper.tex
@@
 \usepackage{pdfpages}
+\usepackage{caption}
 \usepackage{float, algorithmic, algorithm2e, program}
+% \usepackage{algorithm, algpseudocode}
+
 \usepackage{tabularx}
@@
 \newtheorem{definition}{Definition}[section]
+\newlength\myindent
+\setlength\myindent{2em}
+\newcommand\bindent{%
+  \begingroup
+  \setlength{\itemindent}{\myindent}
+  \addtolength{\algorithmicindent}{\myindent}
+}
+\newcommand\eindent{\endgroup}
 
@@
 The reason for doing so is that we only want physically plausible states.
-That is we don't want disconnected states or states that cannot be reached by the system.
+That is, we don't want disconnected states or states that cannot be reached by the system.
 
@@
 It starts with a given starting state in the processing list, then it computes all the successors of this state and adds all the successors as children to state that is being processed, as well as adding these successors to the processing list, but only if a successor was not already processed.
+After the the current state, the next state in the processing list is processed repeating this process iteratively until there are no new states in the processing list.
 
@@
 
-Algorithm \ref{alg:propogating_values} propagates over all quantities and modify value based on the derivative.
-This is done by looping over all quantities in the model and computing all the possible values for a next state based on its current value and its derivative.
+Algorithm \ref{alg:propogating_values} propagates over all quantities and computes new values based on the derivative.
 For each possible next value we create a new state, thus possibly creating multiple states.
-Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created previously.
+Because, we can have new states created by value splits of previous quantities, we need to create new states for each state that was created by a previous split.
 
@@
 Then if both derivatives are possible this implies ambiguity and the 0 derivative is also a possibility.
-If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences).
+If the values of the influences in the previous state did not change this means that the derivative also does not change (w.r.t.\ the influences), this adds partial inequality constraints.
 Finally, for each state we create multiple new states based on the possible derivatives.
@@
 \begin{algorithm}[H]
-	TODO
+    \For{{\bf function} prop\_proportionals(q, visited\_quantities)} {
+		add q to visited\_quantities\\
+		\ForAll {quantities proportionals to q} {
+			\If {quantity has no influences and proportionalities} {
+				return derivative based on exogenous influence
+			}
+			\ElseIf {quantity has no proportionalities} {
+				return derivative based on derivative(s) computed in previous step
+			}
+			\Else{
+				return prop\_proportionals(quantity)
+			}
+		}
+	}
+	\caption{Recursive function for proportionals propagation}
+	\label{alg:propogating_proportionals_re}
+\end{algorithm}
+
+\vspace{12pt}
+\begin{algorithm}[H]
+    \For{quantity in quantities}{
+
+        \ForAll{new\_states}{
+
+            new\_state = pop(new\_states)\\
+
+			\If {quantity not in visited\_quantities} {
+				derivative = prop\_proportionals(quantity)
+			}
+
+			\If{quantity in visited\_quantities does NOT violate \\epsilon ordering or invalid derivative at end values}{
+				add new derivative to new\_state in new\_states
+			}
+
+        }
+
+    }
 	\caption{Proportionals propagation}
@@
 The last step in the algorithm is computing the new derivatives based on the proportionates.
-Again we loop over all the quantities in the model.
+Again we loop over all the quantities in the model, shown in Algorithm~\ref{alg:propogating_proportionals}.
 Then, for each quantity we check the derivatives of the quantities it is proportional to.
-If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm.
+If a quantity it is proportional to, has itself a quantity it is proportional to, we check these quantities first, this is implemented as a recursive algorithm, shown in Algorithm~\ref{alg:propogating_proportionals_re}.
 If the algorithm reaches a quantity that has not proportionates or influences, we base its derivative on an exogenous influence (random).
@@
 In order to prevent the algorithm from entering an infinite loop (due to cycles in the influences graph), we keep track of the quantities we visited and stop visiting them again.
+Lastly the algorithm checks for both continuity of derivatives and epsilon ordering.
 
@@
 All the propagations steps together leave us with possible next states.
-Which are passed to the iterative algorithm for construction a state-graph (Algorithm~\ref{alg:computing_state_graph}).
+Which are passed to the iterative algorithm for the construction of a state-graph (Algorithm~\ref{alg:computing_state_graph}).
 
@@
 First of all we chose not to implement all the inequality constraints.
-The main reason for not including these is that all rules have to be coded for a general problem and thus coding them takes a discrete amount of time.
+The main reason for not including these is that all rules have to be coded for a general problem and we did not find a general solution to this.
 Therefore we decided to implement only the most important ones, Such as the continuity constraints.
 
-We then decided not to use higher order derivatives.
+We also decided not to use higher order derivatives.
 However because of the generality of the algorithm, we could easily have implemented them adding, for each quantity $Q$, an another quantity $\partial Q$, and imposing constraints.
@@
 
-In figure \ref{state-graph-base} shows all possible behaviors of the system described above.
-Nodes are states and arrows represent transitions from a state to every reachable states starting from that source.
+Figure \ref{fig:state_graph_base} shows all possible behaviors of the system described above.
+Nodes are states and arrows represent transitions from a state to every reachable state.
 Nodes are labeled with the description of the current value of each quantity and its derivative.
 
-\begin{figure}[H]\label{state-graph-base}
+\begin{figure}[H]
 \centering
@@
 \caption{Plot of the state-graph for the base problem}
+\label{fig:state_graph_base}
 \end{figure}
@@
 \section{Trace}
-In order to debug the code and let users to better understand what the algorithm do we print a trace.
-The trace of the base problem is in \texttt{base\_trace} file and the trace of the extra problem is in \texttt{extra\_trace} file.
+In order to debug the code as well as creating the possibility for users to better understand what the algorithm does, we generate a trace for the algorithm.
+The trace of the base problem is in \texttt{base\_trace.log} file and the trace of the extra problem is in \texttt{extra\_trace.log} file.
 
 The trace reports how the algorithm expands the state graph and how it determines all the possible transitions for a state.
-
-In particular the trace gives insights on the procedure for computing the possible successors.
+In particular, the trace gives insights on the procedure for computing the possible successors.
 It shows how the values are assigned and how the constraints are propagated.
-It also tells when we discard a partial assignment because it violates value or continuity constraints.
-
-However we decided to show the big picture of what the algorithm is doing than to print every single step in order keep the file readable.
-Thus we don't report all the intermediate steps and computations the programs is performing highlighting only the principal steps.
+It also tells when we discard a partial assignment because it violates any constraints.
 
 \section{Extra Problem}
-
+For the extra problem we use the following quantities and dependencies, and we get a state graph as shown in Figure~\ref{fig:state_graph_extra}.
 \begin{itemize}
@@
 
-\begin{figure}[H]\label{state-graph-extra}
+\begin{figure}[H]
 	\centering
@@
 	\caption{Plot of the state-graph for the extra problem}
+	\label{fig:state_graph_extra}
 \end{figure}
